GPU-Beschleunigung der Material Point Method

### 2019-01-27 mscthesis.tex
final fill things

fill with nonsense

@@

Therefore they propose to use the Fixed Corotated Hyperelasticity energy density:
\begin{equation}
  \Psi_{FCH} = \mu \|\boldsymbol{\Sigma}_E - \boldsymbol{I}\|^2_F + \frac{\lambda}{2}(J_E-1)^2.
\end{equation}
This leads to the primary contour $J=1$ which does not cross the inverted region as only one singular value may be negative.
\begin{flushright}\cite{MPM:INVERT}\end{flushright}
\subsection{Governing equations}
fill with nonsense
deriving the weak form
\subsubsection{Conservation of mass}
Let the Eulerian mass density be $_t\rho(_t\boldsymbol{x},t)$. Similarly, let its (Lagrangian) pull back be $_0\rho(_0\boldsymbol{x},t)$.
After \ref{eq:volume_integral} they are related as:
\begin{equation} \label{eq:density_pull_back}
  \int _ {_tB _ { \epsilon }}_t\rho(_t\boldsymbol{x},t)d_t\boldsymbol{x} =   \int _ {_0B _ { \epsilon }} {^t_0J}_0\rho(_0\boldsymbol{x},t) d_0\boldsymbol{x}.
\end{equation}
An open ball $ _ { 0 }B _ { \epsilon } $ in the reference configuration will have the same mass as its respective open ball $B_{\epsilon}^t$ in the current configuration. Keep in mind that both refer to a deformed state.

@@

\cite{strang2007computational}
\cite{bathe2006finite}

@@

\cite{MPM:APIC}\cite{MPM:COURSE}

@@

\subsection{General-purpose computing on graphics processing units}
fill with nonsense

@@

\clearpage
\section{Evaluation}\label{sec:eval}
fill with nonsense
\subsection{Verifying Results}
\textbf{Parallel reduction}, \textbf{map}, and \textbf{scan} verify against \texttt{std::transform\_reduce} as it supplies the same functionality as a fully reduced parallel reduction. It is standard since C++17. These shaders are exact on integers. Floating point precision can be seen in table \ref{tab:parallel_red}.
\begin{table}[!b]
  \begin{tabular}{|l|r|r|r|}\hline
    Shader       & \texttt{unary\_op} & Total abs. error & Total rel. error \\\hline
    Map (float)   & x          & 0.0    & 0.0      \\\hline
    Map (vec4,dvec4)   & length(x)  &0.0625 & 7.94e-08 \\\hline\hline
    MapReduce (float)& x  &0.0183 & 3.51e-08 \\\hline
    MapReduce (double)& x  &1.78e-08 & 3.42e-14 \\\hline
    MapReduce (vec4)& length(x)  &212.813 & 2.71e-04 \\\hline
    MapReduce (dvec4)& length(x)  &1.58e-05 & 2.02e-11 \\\hline
  \end{tabular}
  \caption{Error of various shaders on 1024*1024 random elements. Scalars are just between [0.0-1.0]. The vectors have length between [0.0-1.0].}
  \label{tab:parallel_red}
\end{table}

@@

\subsection{Performance}
The given metrics in \ref{sec:metrics} lead to valuable insight into how well the GPU can utilize a given shader or program. The GPU was a black box for a long time in this regard. The metrics provided by NVIDIA are a first step to open up the GPUs inner workings. Theory about optimization for the architecture such as \cite{NVIDIA:BEST:PRACTICE} can be more readily applied. However for abstraction purposes theory as well as descriptions about the architecture might be deliberately vague. These metrics give the programmer more power of identifying the right optimizations for a given code.

>>>>

The first chapters will derive the Material-Point-Method and necessary extensions for an elastic model. An elastic model is an interesting case displaying much of the needed computations for a typical model. The last chapter \ref{sec:gpgpu} tends to the GPGPU side.

@@

Therefore they propose to use the Fixed Corotated Hyperelasticity energy density:
\begin{equation}
  \Psi_{FCH} = \mu \|\boldsymbol{\Sigma}_E - \boldsymbol{I}\|^2_F + \frac{\lambda}{2}(J_E-1)^2.
\end{equation}
This leads to the primary contour $J=1$ which does not cross the inverted region as only one singular value may be negative.
\begin{flushright}\cite{MPM:INVERT}\end{flushright}
\subsection{Governing equations}
Starting from the governing equations this chapter will lead to the weak formulation core of any finite element and material point method.
\subsubsection{Conservation of mass}
Let the Eulerian mass density be $_t\rho(_t\boldsymbol{x},t)$. Similarly, let its (Lagrangian) pull back be $_0\rho(_0\boldsymbol{x},t)$.
After \ref{eq:volume_integral} they are related as:
\begin{equation} \label{eq:density_pull_back}
  \int _ {_tB _ { \epsilon }}_t\rho(_t\boldsymbol{x},t)d_t\boldsymbol{x} =   \int _ {_0B _ { \epsilon }} {^t_0J}_0\rho(_0\boldsymbol{x},t) d_0\boldsymbol{x}.
\end{equation}
An open ball $ _ { 0 }B _ { \epsilon } $ in the reference configuration will have the same mass as its respective open ball $B_{\epsilon}^t$ in the current configuration. Keep in mind that both refer to a deformed state.

@@

\begin{flushright}\cite{MPM:COURSE}\cite{strang2007computational}\cite{bathe2006finite}\end{flushright}

@@

\begin{flushright}\cite{MPM:APIC}\cite{MPM:COURSE}\end{flushright}

@@

\subsection{General-purpose computing on graphics processing units}\label{sec:gpgpu}
This chapter is split into two parts. \ref{sec:optim} tends to the memory hierarchy of the GPU and typical optimizations done in GPGPU work. Furthermore, chapter \ref{sec:metrics} will elaborate metrics to quantify optimizations.

@@

\clearpage
\section{Evaluation}\label{sec:eval}
An important part of any operation is to verify them. Chapter \ref{sec:verify} does that. Following up, chapter \ref{sec:perf} will use the metrics out of chapter \ref{sec:metrics} to evaluate the performance of the implementation.
\subsection{Verifying Results}\label{sec:verify}
\textbf{Parallel reduction}, \textbf{map}, and \textbf{scan} verify against \texttt{std::transform\_reduce} as it supplies the same functionality as a fully reduced parallel reduction. It is standard since C++17. These shaders are exact on integers. Floating point precision can be seen in table \ref{tab:parallel_red}.
\begin{table}[!b]
  \begin{tabular}{|l|r|r|r|}\hline
    Shader       & \texttt{unary\_op} & Total abs. error & Total rel. error \\\hline
    Map (float)   & x          & 0.0    & 0.0      \\\hline
    Map (vec4,dvec4)   & length(x)  &0.0625 & 7.94e-08 \\\hline\hline
    MapReduce (float)& x  &0.0183 & 3.51e-08 \\\hline
    MapReduce (double)& x  &1.78e-08 & 3.42e-14 \\\hline
    MapReduce (vec4)& length(x)  &212.813 & 2.71e-04 \\\hline
    MapReduce (dvec4)& length(x)  &1.58e-05 & 2.02e-11 \\\hline
  \end{tabular}
  \caption{Error of various shaders on 1024*1024 random elements. Scalars are just between [0.0-1.0]. The vectors have length between [0.0-1.0].}
  \label{tab:parallel_red}
\end{table}

@@

\subsection{Performance}\ref{sec:perf}
The given metrics in \ref{sec:metrics} lead to valuable insight into how well the GPU can utilize a given shader or program. The GPU was a black box for a long time in this regard. The metrics provided by NVIDIA are a first step to open up the GPUs inner workings. Theory about optimization for the architecture such as \cite{NVIDIA:BEST:PRACTICE} can be more readily applied. However for abstraction purposes theory as well as descriptions about the architecture might be deliberately vague. These metrics give the programmer more power of identifying the right optimizations for a given code.
